A Switching Planner for Combined Task and Observation Planning

### 2011-05-01 code/subarchitectures/planner.sa/trunk/src/latex/AAAI11/evaluation.tex
bunch of changes to smaller details.

We evaluate our approaches in robot exploration tasks from home and
office environments. Spatially, these consist of {\em rooms}
(office/kitchen/etc), and an underlying topological map over
smaller areas of space, called {\em places}, and connectivity between
those. The mobile {\em robot} and {\em visual objects} inhabit the
topological places. Objects indicate the category of space they
inhabit---e.g., spoons are likely to be in kitchens. By examining
view cones at places for particular objects, the robot is able to: (1)
categorise space at high (room) and low (place) levels, and (2) find
objects for the user, exploiting information about object
co-occurrence and room categories for efficiency. Also, in the
presence of a person, the robot can ask about the category of the
current room.

We compare switching to the baseline in several realistic
tasks, with the number of rooms ranging from 3 (12-places, 16-objects,
$|$states$|$ $>10^{21}$) to 6 (26-places, 21-objects, $|$states$|$
$>10^{36}$). We also compare those systems with near optimal policies
computed using Smith's {\sc zmdp} for small 2 room problems (4-places,
3-objects, $|$states$|$ $\simeq 5000$). Our evaluation considers 3
levels of reliability in sensing: {\em reliable} sensors have a $.1$
probability of a false negative, {\em semi-reliable} have a chance of
$0.3$ of false negative and $0.1$ of false positive, and {\em noisy}
sensors with probabilities of $0.5$ and $0.2$ respectively. Each
object class is assigned one sensor model---e.g. cornflakes may be
harder to detect than refrigerators. We performed several experiments
with different levels of reliability for sensing the target object(s),
while keeping sensing models for non-target objects constant.

>>>>

We evaluate our approaches in robot exploration tasks from home and
office environments. Spatially, these consist of {\em rooms}
(office/kitchen/etc), and an underlying topological map over
smaller areas of space, called {\em places}, and connectivity between
those. The mobile {\em robot} and {\em visual objects} inhabit the
topological places. Objects indicate the category of space they
inhabit -- e.g., spoons are likely to be in kitchens. By examining
view cones at places for particular objects, the robot is able to: (1)
categorise space at high (room) and low (place) levels, and (2) find
objects for the user, exploiting information about object
co-occurrence and room categories for efficiency. Also, in the
presence of a person, the robot can ask about the category of the
current room.

We compare switching to the baseline in several realistic
tasks, with the number of rooms ranging from 3 (12-places, 16-objects,
$|$states$|$ $>10^{21}$) to 6 (26-places, 21-objects, $|$states$|$
$>10^{36}$). We also compare those systems with near optimal policies
computed using Smith's {\sc zmdp} for small 2 room problems (4-places,
3-objects, $|$states$|$ $\simeq 5000$). Our evaluation considers 3
levels of reliability in sensing: {\em reliable} sensors have a $.1$
probability of a false negative, {\em semi-reliable} have a chance of
$0.3$ of false negative and $0.1$ of false positive, and {\em noisy}
sensors with probabilities of $0.5$ and $0.2$ respectively. Each
object class is assigned one sensor model -- e.g. cornflakes may be
harder to detect than refrigerators. We performed several experiments
with different levels of reliability for sensing the target object(s),
while keeping sensing models for non-target objects constant.

### 2011-05-01 code/subarchitectures/planner.sa/trunk/src/latex/AAAI11/prop-dtp.tex
bunch of changes to smaller details.

We describe the partially observable propositional probabilistic
planning problem, with costs and rewards. We model a process state
$\state$ as the set of propositions that are true of the
state. Notationally, $\props$ is the set of propositions, $\prop$ is
an element from that set, and we have $\state \subseteq
\props$. The underlying process
dynamics are modelled in terms of a finite set of {\em probabilistic
STRIPS operators}~\cite{boutilier:abstraction} $\stochActions$ over
state-characterising propositions $\propositions$.
%%
We say an action $\stochAction \in \stochActions$ is
applicable if its precondition $\poss(\stochAction)$, a set of
propositions, are satisfied in the current state -- i.e., $\poss(\stochAction) \subseteq
\state$. We denote by $\mu_{\stochAction}(\detAction_i)$ the probability that
nature chooses a deterministic STRIPS effect $\detAct_i$, and for
all \stochAction\ we require
$\sum_{\detAction_i \in \detActions(\stochAction)}
\mu_{\stochAction}(\detAction_i) = 1$.

We are concerned with problems that feature partial
observability. Although we could invoke {\em extended probabilistic
STRIPS operators}~\cite{rintanen:01} to model actions and observations
propositionally, we find it convenient for presentation and computation to
separate sensing and action. Therefore, we suppose a POMDP has a
perceptual model given in terms of a finite set of stochastic {\em
senses} $\stochSenses$, deterministic sensing outcomes $\detSenses$,
and perceptual propositions $\percepts$, called {\em percepts}. In
detail, we take an observation $\observ$ to be a set of percepts
$\observ \subseteq \percepts$, and denote by \observations\ the set of
reachable observations. The underlying state of the process cannot be
observed directly, rather, senses $\stochSense \in \stochSenses$
effect an observation $\observ \in
\observations$ that informs what should be believed about the state the
process is in. If $\action$ is applied effecting a
transition to a successor state $\state'$, then an observation occurs
according to the active senses $\stochSenses(\stochAction, \state')
\subseteq \stochSenses$. A sense $\stochSense$ is active, written
$\stochSense \in \stochSenses(\stochAction, \state')$, if the senses'
action-precondition, $\poss_\stochActions(\stochSense)$, is equal to
$\stochAction$, and the state-precondition $\poss_\states(\stochSense)
\subseteq \propositions$ is satisfied by the state $\state'$ -- In the
usual sense that \mbox{$\poss_\states(\stochSense)\subseteq\state'$}.
%%
When a sense is active, nature must choose exactly one outcome amongst
a small set of deterministic choices $\detSenses(\stochSense)
\equiv \{\detSense_1, \ldots, \detSense_k \}$, so that for each
$i$ we have $\detSense_i \subseteq \percepts$. The probability of
the $i^{th}$ element being chosen is given by
$\psi_{\stochSense}(\detSense_i)$, where $\sum_{\detSense_i \in
\detSenses(\stochSense)} \psi_{\stochSense}(\detSense_i) =
1$. The observation received by the agent corresponds to the union of
perceptual propositions from the chosen elements of active
senses.

@@

Until now we have discussed the POMDP in terms of propositions and
percepts. In order to address belief revision and utility it is
convenient to consider the underlying decision process in a flat
format. This is given by the tuple
$\langle \states, \bstate_0, \actions, \transProb, \reward,
\observations, \obsDist \rangle$. Here, $\bstate_0$ is the initial
belief-state, \states\ is the finite set of reachable propositional
states, \actions\ is the finite set of actions, and \observations\ is
the finite set of reachable observations (i.e., perceptual states).
Where $\state, \state' \in \states$, $a \in \actions$, from $\mu$ we
have a state transition function $\transProb(\state, \action,
\state')$ giving the probability of a transition from state $s$ to
$s'$ if $a$ is applied. For any $\state$ and $\action$ we have
$\sum_{\state' \in \states} \transProb(\state, \action, \state') = 1$.
%%
The function $\reward:\states \times \actions \to \Re$ is a bounded real
valued reward function. Therefore a finite positive constant $c$
exists so that for all $\state \in \states$ and $\action \in
\actions$, $|\reward(\state, \action)| < c$. We model costs as
negative rewards.
%%
From $\psi$ we have that for each $\state' \in \states$ and action
$\action \in \actions$, an observation $\observ \in \observations$ is
generated independently according to a probability distribution
$\obsDist(\state', \action)$. We denote by $\obsDist_\observ(\state',
\action)$ the probability of getting observation $\observ$ in state
$\state'$. For $\state'$ and $\action$ we have $\sum_{\observ \in
\observations} \obsDist_\observ(\state', \action) = 1$.

@@

\noindent where $\pp{Pr}(\observ | \action, \bstate)$ is the
normalising factor, that is, the probability of getting observation
$\observ$ given we execute $\action$ in $\bstate$.

>>>>

We describe the partially observable propositional probabilistic
planning problem, with costs and rewards. We model a process state
$\state$ as the set of propositions that are true of the
state. Notationally, $\props$ is the set of propositions, $\prop$ is
an element from that set, and we have $\state \subseteq
\props$. The underlying process
dynamics are modelled in terms of a finite set of {\em probabilistic
STRIPS operators}~\cite{boutilier:abstraction} $\stochActions$ over
state-characterising propositions $\propositions$.
%%
We say an action $\stochAction \in \stochActions$ is
applicable if its precondition $\poss(\stochAction)$, a set of
propositions, are satisfied in the current state -- i.e., $\poss(\stochAction) \subseteq
\state$. We denote by $\mu_{\stochAction}(\detAction_i)$ the probability that
nature chooses a deterministic STRIPS effect $\detAct_i$, and for
all \stochAction\ we require
$\sum_{\detAction_i}
\mu_{\stochAction}(\detAction_i) = 1$.

We are concerned with problems that feature partial
observability. Although we could invoke {\em extended probabilistic
STRIPS operators}~\cite{rintanen:01} to model actions and observations
propositionally, we find it convenient for presentation and
computation to separate sensing and action. Therefore, we suppose a
POMDP has a perceptual model given in terms of a finite set of
stochastic {\em senses} $\stochSenses$, deterministic sensing outcomes
$\detSenses$, and perceptual propositions $\percepts$, called {\em
percepts}. In detail, we take an observation $\observ$ to be a set of
percepts , i.e., $\observ \subseteq \percepts$, and denote
by \observations\ the set of \Omit{reachable} observations. The underlying
state of the process cannot be observed directly, rather, senses
$\stochSense \in \stochSenses$ effect an observation $\observ \in
\observations$ that informs what should be believed about the state the
process is in. If action $\action$ is applied effecting a
transition to a successor state $\state'$, then an observation occurs
according to the active senses $\stochSenses(\stochAction, \state')
\subseteq \stochSenses$. A sense $\stochSense$ is active, written
$\stochSense \in \stochSenses(\stochAction, \state')$, if the senses'
action-precondition, $\poss_\stochActions(\stochSense)$, is equal to
$\stochAction$, and the state-precondition $\poss_\states(\stochSense)
\subseteq \propositions$ is satisfied by the state $\state'$, i.e., \mbox{$\poss_\states(\stochSense)\subseteq\state'$}.
%%
When a sense is active, nature must choose exactly one outcome amongst
a small set of deterministic choices $\detSenses(\stochSense)
\equiv \{\detSense_1, \ldots, \detSense_k \}$, so that for each
$i$ we have $\detSense_i \subseteq \percepts$. The probability of
the $i^{th}$ element being chosen is given by
$\psi_{\stochSense}(\detSense_i)$, where $\sum_{\detSense_i \in
\detSenses(\stochSense)} \psi_{\stochSense}(\detSense_i) =
1$. The observation received by the agent corresponds to the union of
perceptual propositions from the chosen elements of active
senses.

@@

Until now we have discussed the POMDP in terms of propositions and
percepts. In order to address belief revision and utility it is
convenient to consider the underlying decision process in a flat
format. This is given by the tuple
$\langle \states, \bstate_0, \actions, \transProb, \reward,
\observations, \obsDist \rangle$. Here, $\bstate_0$ is the initial
belief-state, \states\ is the finite set of reachable propositional
states, \actions\ is the finite set of actions, and \observations\ is
the finite set of reachable observations.  Where
$\state, \state' \in \states$, $a \in \actions$, from $\mu$ we have a
state transition function $\transProb(\state, \action,
\state')$ giving the probability of a transition from state $s$ to
$s'$ if $a$ is applied. For any $\state$ and $\action$ we have
$\sum_{\state' \in \states} \transProb(\state, \action, \state') = 1$.
%%
The function $\reward:\states \times \actions \to \Re$ is a bounded real
valued reward function. Therefore a finite positive constant $c$
exists so that for all $\state \in \states$ and $\action \in
\actions$, $|\reward(\state, \action)| < c$. We model costs as
negative rewards.
%%
From $\psi$ we have that for each $\state' \in \states$ and action
$\action \in \actions$, an observation $\observ \in \observations$ is
generated independently according to a probability distribution
$\obsDist(\state', \action)$. We denote by $\obsDist_\observ(\state',
\action)$ the probability of getting observation $\observ$ in state
$\state'$. For $\state'$ and $\action$ we have $\sum_{\observ \in
\observations} \obsDist_\observ(\state', \action) = 1$.

@@

\noindent where $\pp{Pr}(\observ | \action, \bstate)$ is a
normalising factor, giving the probability of getting observation
$\observ$ if $\action$ is applied to $\bstate$.

### 2011-05-01 code/subarchitectures/planner.sa/trunk/src/latex/AAAI11/relatex.tex
bunch of changes to smaller details.

There has also been much recent work on scaling offline approximate
POMDP solution procedures to medium-sized instances. Recent
contributions propose more efficient belief-point sampling
schemes~\cite{kurniawati:etal:2010,shani:etal:08}, and factored
representations with procedures that can efficiently exploit
structures in those
representations~\cite{brunskill:russell:2010,shani:etal:2008}. Offline
domain independent systems scale to {\em logistics} problems with
$2^{22}$ states~\cite{shani:etal:2008}, taking over an hour to
converge, and around 10 seconds on average to perform each Bellman
backup. \citeauthor{brunskill:russell:2010} are able to solve problems
with approximately $10^{30}$ states, by further exploiting certain
problem features -- E.g., problems where no actions have negative
effects.
%%
Moving someway towards supporting real-time decision making, recent
online POMDP solution procedures have been developed which leverage
highly approximate value functions -- computed using an offline
procedure -- and heuristics in forward
search~\cite{ross:etal:2008}. These approaches are applicable in
relatively small problems, and can require
expensive \emph{problem-specific} offline processing in order to yield
good behaviours.
%%
A {\em very} recent and promising online approach for
larger POMDPs employs Monte-Carlo sampling to break the curse of
dimensionality in situations where goal reachability is {\em easily}
determined~\cite{silver:veness:2010}.

In the direction of leveraging {\em classical} systems/approaches for
planning under uncertainty, the most highlighted system to date has
been \system{FFR$_a$}~\cite{yoon:etal:2007}; The winning entry from
the probabilistic track of the 2004 International Planning
Competition.  In the continual paradigm, \system{FFR$_a$}
uses \system{FF} to compute sequential plans and execution traces.
%%
More computationally expensive approaches in this vein combine
sampling strategies on valuations over {\em runtime variables} with
deterministic planning procedures~\cite{yoon:etal:2008}. %% The outcome is typically a more
%% robust sequential plan, or contingent
%% plan~\cite{majercik:2006}.

Also leveraging deterministic planners in problems that feature
uncertainty, \system{Conformant-FF}~\cite{hoffmann:brafman:2006} and
$T_0$~\cite{palacios:geffner:2009} demonstrate how conformant planning
---i.e., sequential planning in unobservable worlds--- can be modelled
as a deterministic problem, and therefore solved using sequential
systems. In this conformant setting, advances have been towards
compact representations of beliefs amenable to existing best-first
search planning procedures, and lazy evaluations of beliefs. Most
recently this research thread has been extended to contingent planning
in fully observable non-deterministic
environments~\cite{albore:etal:2009}.

%% Their applicability in our setting is
%% limited, firstly because they only scale to smaller problems, with
%% thousands of states, and also due to the large amount of
%% \emph{problem-specific} offline processing that might be required to get useful
%% search guidance.

%%  In that
%% setting \citeauthor{kurniawati:etal:2010}~(\citeyear{})
%% recently addressed an inefficiency of offline point-based techniques
%% in problems with medium length planning horizons, however their
%% approach take tens of minutes to good plans, and only scales to tens
%% of thousands of states. In the case of general domain-independent
%% factored systems, the state-of-the-art scales to relatively small
%% problems with $2^{22}$ states~\cite{shani:etal:2008}.

%% At their limit,
%% these procedures take over an hour to converge, and $\sim10$ seconds
%% on average to perform a single Bellman backup.  For classes of POMDP
%% that feature exploitable structures (e.g., no actions with negative
%% effects), problems with as many as $10^{30}$ states can be targeted by
%% such offline procedures~\cite{brunskill:russell:2010}.

%% Although we suppose it an interesting
%% item for future work to pursue that direction, it should be noted that
%% ease of goal reachability is not guaranteed in the problems we face,
%% and is certainly not a property to be assumed in domain independent
%% planning.

%%
%% The continual planning system that motivated our
%% project~\cite{brenner:nebel:jaamas09} also has this characteristic,
%% and has been applied in completely observable domains, particularly
%% those featuring multiple communicating agents.

%% The use of knowledge
%% operators in domains allows plans that act to gain knowledge, but the
%% approach assumes that such actions are deterministic and reliable, an
%% assumption that we relax.

%%No! They simply haven't been evaluated in PO settings. They may, or
%%may not struggle. They have sampling of traces, and that would
%%include observations, and therefore evolutions of beliefs. SSAT was
%%proposed by Littman for POMDPs. So the majercik stuff is perfectly
%%suited to POMDPs.

% Normally if we are going to compare with related work, we do
% actually *compare*. Why didn't you try those approaches? I think
% it's safe to say that FFR will struggle. Why would it even include
% in its plan an observational action that doesn't change the world?

%%
%% However, as we said in the introduction,
%% all these approaches struggle in partially observable domains as they
%% rely on being able to determine the state at all times.

>>>>

There has also been much recent work on scaling offline approximate
POMDP solution procedures to medium-sized instances. Recent
contributions propose more efficient belief-point sampling
schemes~\cite{kurniawati:etal:2010,shani:etal:08}, and factored
representations with procedures that can efficiently exploit
structures in those
representations~\cite{brunskill:russell:2010,shani:etal:2008}. Offline
domain independent systems scale to {\em logistics} problems with
$2^{22}$ states~\cite{shani:etal:2008}, taking over an hour to
converge, and around 10 seconds on average to perform each Bellman
backup. \citeauthor{brunskill:russell:2010} are able to solve problems
with approximately $10^{30}$ states, by further exploiting certain
problem features -- e.g., problems where no actions have negative
effects.
%%
Moving someway towards supporting real-time decision making, recent
online POMDP solution procedures have been developed which leverage
highly approximate value functions -- computed using an offline
procedure -- and heuristics in forward
search~\cite{ross:etal:2008}. These approaches are applicable in
relatively small problems, and can require
expensive \emph{problem-specific} offline processing in order to yield
good behaviours.
%%
A {\em very} recent and promising online approach for
larger POMDPs employs Monte-Carlo sampling to break the curse of
dimensionality in situations where goal reachability is {\em easily}
determined~\cite{silver:veness:2010}. 
%%
Our approach can also be thought of as an online POMDP solver that
uses a sequential plan to guide the search, rather than (e.g.,
Monte-Carlo) sampling. Also, compared to most online POMDP procedures, which
replan at each step, our approach involves relatively little replanning.

%% Our approach requires minimal DT replanning,
%% preferring

%% Also, our approach will execute a few
%% actions (as much as it can before an assumption is found to be true or
%% false) before needing to plan again, rather than planning after each
%% step.

%% Moving someway towards supporting real-time decision making, recent
%% online POMDP solution procedures have been developed which leverage
%% highly approximate value functions -- computed using an offline
%% procedure -- and heuristics in forward
%% search~\cite{ross:etal:2008}. These approaches are applicable in
%% larger problems (\cite{ross:etal:2008} includes experiments in a domain
%% with around 12 thousand states), but the fact they often require
%% expensive \emph{problem-specific} offline processing in order to yield
%% good behaviours means they are impractical in problems as large as ours.
%% %%
%% A recent and promising online approach for
%% larger POMDPs employs Monte-Carlo sampling to break the curse of
%% dimensionality in situations where goal reachability is {\em easily}
%% determined~\cite{silver:veness:2010}. Our approach can also be thought of
%% as an online POMDP solver, but uses the sequential plan to guide the
%% search
%% for a decision-theoretic plan, rather than Monte-Carlo sampling. In
%% addition, our approach will execute a few actions (as much as it can
%% before
%% an assumption is found to be true or false) before needing to plan again,
%% rather than planning after each step.

%% [have a look at the bit in parentheses above, I'm not sure I quite
%% characterised how much plan it executes before replanning satisfactorally]

In the direction of leveraging {\em classical} approaches for
planning under uncertainty, the most highlighted system to date has
been \system{FFR$_a$}~\cite{yoon:etal:2007}; The winning entry from
the probabilistic track of the 2004 International Planning
Competition.  In the continual paradigm, \system{FFR$_a$}
uses \system{FF} to compute sequential plans and execution traces.
%%
More computationally expensive approaches in this vein combine
sampling strategies on valuations over {\em runtime variables} with
deterministic planning procedures~\cite{yoon:etal:2008}. %% The outcome is typically a more
%% robust sequential plan, or contingent
%% plan~\cite{majercik:2006}.

Also leveraging deterministic planners in problems that feature
uncertainty, \system{Conformant-FF}~\cite{hoffmann:brafman:2006} and
$T_0$~\cite{palacios:geffner:2009} demonstrate how conformant planning
-- i.e., sequential planning in unobservable worlds -- can be modelled
as a deterministic problem, and therefore solved using sequential
systems. In this conformant setting, advances have been towards
compact representations of beliefs amenable to existing best-first
search planning procedures, and lazy evaluations of beliefs. We
consider it an appealing future direction to pursue conformant
reasoning during the sequential sessions we proposed. Most recently
this research thread has been extended to contingent planning in fully
observable non-deterministic environments~\cite{albore:etal:2009}.
